Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-07-27 cover_letter/response_letter_MJ_210721.tex

\noteg{Can also include a figure for the F4 or shifted states showing a minimal difference.}

@@

\textcolor{blue}{8. I don't understand how the population of F4 could be >50\% that of H for the GC-core system, when the free energy of F4 is reported as > 5kJ/mol. My back-of-the-envelope maths predicts a ratio of 8:1 for a 5kJ/mol gap. Incidentally, 8:1 is roughly the ratio I got from running the configurations through NUPACK.}

\textbf{Author Reply:}  \noteg{I added everything below, not sure if this is adequately adressing the mistake} We apologize for the confusion and have found that it stems from a unit conversion error for the GC-core sequence. Specifically, we did not convert the $\Delta$ F (5.1 kT) from units kT to kJ/mol, which should result in a much larger $\Delta$ F (13.7 kJ/mol) relative to the hybridized state.

We recognize that this is a significant miscalculation, and we have re-checked all free energy results to ensure it was an error for the GC-core free energy calculations only and did not affect the reported values for GC-end and AT-all. Furthermore, we have verified that the error does not propagate to any other relevant results in the manuscript.

We have updated Figure 3 with the correct $\Delta$ F  as shown below.

\begin{figure}[ht!]
	\begin{center} 
        \includegraphics[width=160mm, scale=1]{Fig3.pdf}
        \caption{Comparison of the macrostate free energy predictions of the MSMs and nearest neighbor (NN) thermodynamic model \rood{at the sequence melting temperatures}.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} (a) Free energies of each macrostate relative to the hybridized state $\Delta F = F - F_H$. We define the hybridized state H to possess a free energy of zero and take care to only compare relative free energies (i.e., $\Delta F$) between the MSM and NN model. (b) Discrepancy between the macrostate relative free energy predictions $\Delta \Delta F = \Delta F^\mathrm{MSM} - \Delta F^\mathrm{NN}$ of the MSM relative to the NN model. The MSM tends to predict higher relative free energies (i.e., lower occupancy probabilities) for the out-of-register shifted states 5S2, 3S2, 5S4, 3S4 relative to the hybridized state H compared to the NN model.}
        \label{fig:NN_table}
	\end{center}
\end{figure}

We have also updated the second paragraph of section 3.3 as follows:

As illustrated in Fig.~\ref{fig:NN_table}, we see that the MSM tends to predict higher free energies for \rood{all macrostates relative to the H state compared to the NN model}.

\noteg{The line above is replacing the following: ``As illustrated in Fig.~\ref{fig:NN_table}, we see that the MSM tends to predict higher free energies for the out-of-register shifted macrostates 5S2, 3S2, 5S4, and 3S4 relative to the H state compared to the NN model. The NN and MSM predictions for the frayed state F4 are in good agreement.''}

\noteg{Do we still want to include these calculations or just say that it was wrong?}

Our back-of-the-envelope calculations yield a similar result based on the $\Delta F$ values reported in Fig.~3. In brief, assuming:

@@

We believe that the root of this comment may be our choice of a logarithmic y-axis for the bar plots in Fig.~2b that make the bar associated with the F4 state appear nearly 50\% of the height of the H and D bars for GC-core. We made this plotting choice in order to make visible on these bar plots the small state populations associated with the shifted and frayed states and noted the percentage occupancies on the bars themselves. To help eliminate this confusion, we have modified the portion of the Fig.~2 caption referring to this panel to read:

``(b) \textbf{Thermodynamics.} Histograms reporting the number of the $10^7$ total frames within the 1 ms of simulation trajectories observed to occupy each of the seven macrostates, corresponding to our numerical estimates of the equilibrium occupancy probabilities. Uncertainties are calculated across 100 MSMs using a Bayesian MSM estimation are reported for each bar and are very small compared to the total counts. \rood{Values are reported on a log y-axis to make the small populations of the shifted and frayed states visible.}''

\textcolor{blue}{9. Is the MSM model explicitly constrained to obey detailed balance? The numbers that come out seem consistent with detailed balance - forward and backward trajectories at the same T essentially look the same (they're equally likely to go through intermediates). This should be a feature of the MSM, but it's unclear to me whether it's hard-coded or just arises naturally in the coarse-graining.}

\textbf{Author Reply:}    The reviewer raises a good point regarding the assumptions made during MSM construction. We do indeed use the reversible Bayesian MSM framework described here \url{http://www.emma-project.org/latest/api/generated/pyemma.msm.BayesianMSM.html} which enforces detailed balance. This is an important point which we have now made explicit by modifying the text in Section 2.1.2 on p.~11 to read:

``The 200 microstates comprising each system were coarsened into our terminal macrostate MSM. \rood{Since we are interested in the construction of models of the equilibrium kinetics, we explicitly enforce detailed balance in the construction of the MSM. \citep{Wehmeyer2019IntroductionSoftware}}''

and the text in Section 3.4 on p.~26 to read:

@@

``In addition to thermodynamic stabilities, the macrostate MSM also furnishes quantitative and interpretable predictions of hybridization and dehybridization pathways and mechanisms. \rood{It should be noted that because we are using a reversible MSM framework, detailed balance is enforced by construction.''}

\textcolor{blue}{Relatedly, some of the discussion of the hybridization and de-hybridization pathways in non-repetitive sequences seems to ignore this fact. In these equilibrium simulations, hybridization trajectories played in reverse should be statistically indistinguishable from de-hybridization trajectories. I don't think anything that's said is wrong, per se, but it might be worth considering how the contrast between "nucleation and zippering" and "fraying-peeling" is consistent with the symmetry in the pathways.}

\textbf{Author Reply:}  This is a nice observation, and we would do well to note this symmetry for the reader. We have added a sentence to this effect in Section 3.6 on p.~31:

``In dehybridization, we observe fraying \rood{of one half of the duplex with the strands remaining associated until the loss of the final A:T pairs and ultimately the last G:C pair}. Qualitatively, we observed some short-lived states composed of two to four native WC base pair contacts immediately before full dissociation occurs, but, in contrast to the F4 state we observe in GC-core, these conformations do not constitute a metastable state within our MSM nor do they tend to reform intact duplexes. These dissociation dynamics are consistent with a ``fraying-peeling'' dehybridization mechanism.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding, Zgarbova2014BaseRNA} \rood{We observe that the principle of microscopic reversibility for a molecular system at thermodynamic equilibrium imposes the conditions of detailed balance and symmetry of the classical equations of motion under time reversibility \cite{McCully2008MicroscopicHomeodomain}. A consequence of these considerations is that if our system is indeed at equilibrium, then the ``fraying-peeling'' dehybridization mechanism can be considered a reversal of the ``nucleation-zippering'' hybridization pathway since the time-reversed simulation trajectories represent an equally valid sampling of the system dynamics. This is indeed borne out by our mechanistic observations for GC-mix wherein the early stages of ``nucleation-zippering'' proceed by the formation of one G:C contact followed by one or more additional A:T pairs and the late stages by the formation of all remaining WC pairs, which we compare with the early stages ``fraying-peeling'' wherein one half of the duplex frays and the late stages wherein dissociation finally completes by the dissolution of the last few A:T contacts and the final G:C pair.}''

@@

\textcolor{blue}{10. I think some of the characterisation of the "all or nothing" picture is a bit misleading. The two state model of duplex formation is really a thermodynamic model, based on the assumption that the ensemble of hybridized and de-hybridized states look very different and are separated by a large barrier. This model is perfectly consistent with the transition between the two ensembles involving a complex network of metastable states. Indeed, as the authors note on line 3 of p4, a number of previous authors have considered what such states might look like, particularly for repetitive sequences. The authors' primary contribution here is not proposing that these states exist for the first time, but trying to tie together evidence from simulation and experiment to describe them.}

\textbf{Author Reply:}   We thank the reviewer for this important observation and illuminating our need to better clarify this potentially ambiguous concept. We have modified the text in Section 3.6 on p.~29 to read:

\noteg{Added sentence starting with ``Indeed, the sequence...'' Not totally sure about this, but I don't think we need to remove all references to all-or-nothing}
``The GC-mix sequence is the only one of the four sequences studied that exhibits simple two-state ``all-or-nothing'' behavior.~\citep{Xiao2019, Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates, Sanstead2016} \rood{This characterization is intended to reflect that the mechanistic observation that} GC-mix MSM comprises just two states, the hybridized H and dehybridized D (Fig.~\ref{fig:allseq_table}c), indicating that association and dissociation of the strands proceeds directly without passing through any metastable intermediate states resolvable under the $\tau$=1.2 ns lag time of our MSM. \rood{Indeed, the sequence's propensity to fray in simulation and the role of termini fraying in dissociation experiments \citep{Sanstead2016} indicate that there remain deviations from ``all-or-nothing'' behavior which are not fully captured within the MSM lag time. (We note that ``all-or-nothing'' can also be used, somewhat ambiguously, to refer to a two-state thermodynamic model where the hybridized and dehybridized are separated by a large free energy barrier but that the transitions between them may proceed through a network of metastable states.)}''

\noteg{Brennan's recommendation:}
We have also removed ``all-or-nothing'' from the experimental methods section as the fitting procedure is more appropriately described as simply ``two-state.''

``FTIR temperature series were measured for each sequence as reported previously,~\citep{Sanstead2016} and the second SVD component along temperature was fit to a \rood{two-state model} to determine the fraction of intact duplex $\theta$ as a function of temperature.''

@@

\textcolor{blue}{11. Related to the above, the references on line 3 of p4 should probably also include Ref 52, where a similar analysis of metastable states in a repetitive sequence is performed (for a different coarse-grained model) and [Flamm, W., Fontana,J.M., Hofacker, I.L. and Schuster,P. (2000) RNA folding at elementary step resolution. RNA, 6, 325-338] where a defect diffusion mechanism is proposed.
}

\textbf{Author Reply:}   We have added this reference.

\textcolor{blue}{12. The reasoning behind why the broader relaxation spectrum is indicative of misaligned states was not that clear to me. Why would "elevated" fraying cause a broader response? Wouldn't it also, then, be broader for CG-core, which has elevated fraying? It seems to me that the argument is more along the lines of the heterogeneity of the configurations leading to heterogeneous dynamics, and hence a broader relaxation profile. But I am no expert on this technique.}

\textbf{Author Reply:}   The reviewer makes a good comment on our choice of language used to described the stretched AT-all response. It is indeed the heterogeneity of the configurations in the initial ensemble, and the fraying dynamics exhibited in each of these states, that is the origin of the broadening of the relaxation profile, but our wording here was insufficiently clear. We believe that the reviewer's apt suggestion ``heterogeneity of configurations leads to heterogeneous dynamics'' expresses this idea much more clearly and we have integrated it into our edits of Section 3.7 on p.~33:

@@

``Nevertheless, the presence of these out-of-register shifted states in the low-temperature equilibrium ensemble prior to the T-jump step may be observable via their influence on the fast T-jump IR response attributable to fraying. Specifically, we hypothesize that the dangling ends and inert tails present in the out-of-register shifted states should promote an \rood{elevated fraying response over the course of the relaxation that is distinct from that of in-register fraying. This heterogeneity of configurations should lead to heterogeneous dynamics, manifested in the observation of a more stretched relaxation over experimental time scales of 70-100 ns.} Analysis of the MSM equilibrium distributions reveals 10.0\% of the equilibrium ensemble to reside in out-of-register shifted states 3S2, 5S2, 3S4, and 5S4 for AT-all, compared to just 0.23\% for GC-end, and 0\% for GC-core and GC-mix. It is our conjecture that a substantial population of out-of-register shifted states in the pre-T-jump AT-all ensemble should be distinguishable from the GC-end, GC-core, and GC-mix as an elongation of the fast relaxation response associated with terminal base fraying.''

\textcolor{blue}{13. How much of the code is being made available? Ideally, the initialization files for the simulations and the code used to perform the MSM analysis would be publicly available. Additionally, it would be good to publish the raw experimental data and the processing tools used to obtain the final results.}

\textbf{Author Reply:}   As a computationally-focused paper, we agree that the data and analysis code should be made available to the reader. We have uploaded input files and analysis scripts to Github (\url{https://github.com/mrjoness/Oligo-MSMs}) and 45 GB of featurized trajectories and experimental data to Zenodo (\url{https://doi.org/10.5281/zenodo.5140040}). The links to both of these resources are now noted in the Supporting Information statement on p.~38:

``\rood{Input files and analysis scripts have been uploaded to Github (\url{https://github.com/mrjoness/Oligo-MSMs}) and 45 GB of featurized trajectories and experimental data to Zenodo (\url{https://doi.org/10.5281/zenodo.5140040}).}''
